\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage{verbatim}
\usepackage{tikz}
\usepackage{hyperref}
\setbeamertemplate{footline}[frame number]
\title{Quick R Tips}
\author{Heide Jackson }
\date{August 2019}

\begin{document}

\maketitle

\begin{frame}{High Level Things to Know}
\begin{itemize}
\item R is free and always will be.
\item It's a flexible statistical software environment based on the S language.
\item R is dynamic and updated more frequently than other statistical software packages like SAS or Stata.
\item R is user driven and its functionality is greatly enhanced by user contributed packages.
\item R Studio is a nice wrapper for accessing R
\end{itemize}
\end{frame}

\begin{frame}[fragile]{The basics of doing anything in R}
\begin{verbatim}
object<-function(...)
or equivalently
object=function()
\end{verbatim}
\begin{itemize}
    \item Functions may be built into R's basic system, loaded in from packages, or defined by you, the user.
\end{itemize}
\end{frame}

\begin{frame}{Other Things To Know}
\begin{itemize}
    \item R is case sensitive. E.g. x and X will be stored in R as two different objects.
    \item The help function is your friend.  If you are trying to figure out what a function does, type help(functioname) or equivalently ?functionname.
    \item Objects can be embedded in other objects.  If you call data\$x, this can be different from calling x.
   \item \# is used prior to a comment.  Multi line comments aren't supported but there are system and program specific shorts cuts for this if you want them.
    \end{itemize}
    \end{frame}


\begin{frame}{Common Hazards and Missteps of Working in R}{Redundancy}
\begin{itemize}

    \item Redundancy is major limitation and benefit to working in R.
    \begin{itemize}
        \item Any task, no matter how simple or mundane can be accomplished at least five different ways using R or user supplied packages.
        \item Users may be overwhelmed by choice. 
        \item The benefit to this is any task can be effectively dual coded.  When in doubt about what a function is doing, look for another package that does the same thing.
        \item When trying to decide how to code, it's sometimes worth doing a task in the easiest way and using base R.  
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Common Hazards and Missteps of Working in R}{Quality Control}
\begin{itemize}
    \item Great R programmers may not be great researchers and vice versa.
    \item How do you evaluate the quality of an R Package?
    \begin{itemize}
        \item For common tasks, packages that have been more extensively used tend to be better debugged.  
        \item Right now the tidyverse collection of packages has  become the R version of mainstream.
        \item For complex methods, research the contributor.  Where have they published?  What documentation do they provide?  What vignettes do they create?  What data do they use?
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Common Hazards and Missteps of Working in R}{Version Control}
\begin{itemize}
    \item R and its packages are frequently updated.  R code may work one day and crash the next.
    \item Package control packages can help resolve this.  A couple of options to consider are docker and packrat.
    \item As a global disclaimer, just because code is reproducible doesn't mean that it is correct.  
    
\end{itemize}
\end{frame}

\end{document}
